---
title: "Dynamical Network Biomarkers"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

An R package to calculate Dynamical Network Biomarkers (DNB).

For more information, see:

[Chen L, Liu R, Liu Z-P, Li M, Aihara K. Detecting early-warning signals for sudden deterioration of complex diseases by dynamical network biomarkers. Scientific Reports. 2012;2:342. doi:10.1038/srep00342](https://www.nature.com/articles/srep00342)

## Install

```{r eval=FALSE}
library(devtools)
install_github("gpli/DNB")
```

## Quick start

```{r quick-start, results='hide', warning=FALSE}
library(DNB)

# Load data
gene_expr <- read.csv(system.file("extdata", "gene_expr.csv", package = "DNB"), row.names = 1)
sample_info <- read.csv(system.file("extdata", "sample_info.csv", package = "DNB"), row.names = 1)

# New DNB object
# all-zero genes at any time point will be removed
dnb <- new_DNB(data = gene_expr,
               time = as.factor(sample_info$time))

# Calulate correlation and coefficient of variation
dnb <- cal_cor(dnb)
dnb <- cal_cv(dnb)

# Search DNB
dnb <- search_candidates(dnb, min_size = 20, max_size = 100)
dnb <- cal_final(dnb)

# Get results
dnb_genes <- get_DNB_genes(dnb)
candidates <- get_candidates(dnb)
plot_DNB(candidates)
final <- get_final(dnb)
plot_DNB(final)
```


## Usage

### Normal samples
Set `group` in `new_DNB` to define normal samples (control group).

```{r eval=FALSE}
dnb <- new_DNB(data = gene_expr,
               time = as.factor(sample_info$time),
               group = as.factor(sample_info$group))
```

By default, control group will be considered during searching DNB, but you can still change your mind by set `with_ctrl = F` in `search_candidates` and `cal_final`. This is useful for tuning parameters.

### Calculating correlation
Calculating correlation are often memory consuming if there are too many genes, and difficult to load them all if there are too many time points. You can set `use_bigcor = T` to calculate correlation with `propagate::bigcor` and/or set `data_dir` to store correlation data to local, i.e.

```{r eval=FALSE}
cal_cor(dnb, use_bigcor = T, data_dir = "tmp")
```

It is recommended to pre-filter low expression genes, which will improve calculation speed.

### Output
`plot_DNB` use `ggplot2` to draw results. You can set `outfile` to write into a image file. Further arguments were passed to `ggplot2::ggsave`.

```{r eval=FALSE}
plot_DNB(final, "DNB_final.pdf")
```
